### 2020-07-05 chapters/background.tex
Abbrevs

Two of the most basic SWAG algorithms are \textit{Recalculate-From-Scratch (RFS)} and \textit{Subtract-On-Evict (SOE)}. Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing. Recall that the combine function is represented by $\oplus$ and the identity or neutral element of the aggregation by $\overline{\theta}$. Pseudocode for the Recalculate-From-Scratch algorithm is shown in Algorithm \ref{pseudo:rfs}.

\begin{algorithm}
    \caption[Recalculate-From-Scratch]{Recalculate-From-Scratch insert, evict and query methods}
    \label{pseudo:rfs}
    \begin{algorithmic}[1]
        \Function{query}{vals}
            \State $agg\gets\overline{\theta}$
            \ForEach {$v \in vals$}
                \State $agg \gets agg \oplus v$
            \EndFor
            \State \Return $agg$
        \EndFunction

        \Function{insert}{vals, v}

            \State $vals.pushBack(v)$
        \EndFunction

        \Function{evict}{vals}

            \State $vals.popFront()$
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element inserted or evicted from the window, the aggregation value is updated without recomputing the entire aggregation like in RFS. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations, that is, a function $\ominus$ exists such that $(x \oplus y) \ominus y = x$, for all $x$ and $y$. To illustrate the differences between SOE and RFS, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the current total sum plus ($\oplus$) the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed thus far minus ($\ominus$) the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable. Algorithm \ref{pseudo:soe} shows pseudocode for SOE.

\begin{algorithm}
    \caption[Subtract-On-Evict]{Subtract-On-Evict insert, evict and query methods}
    \label{pseudo:soe}
    \begin{algorithmic}[1]
        \Function{query}{}

            \State \Return $agg$
        \EndFunction

        \Function{insert}{vals, v}

            \State $vals.pushBack(v)$
            \State $agg \gets agg \oplus v$
        \EndFunction

        \Function{evict}{vals}

            \State $v \gets vals.popFront()$
            \State $agg \gets agg \ominus val$
        \EndFunction
    \end{algorithmic}
\end{algorithm}

>>>>

Two of the most basic SWAG algorithms are Recalculate-From-Scratch (RFS) and Subtract-On-Evict (SOE). Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing. Recall that the combine function is represented by $\oplus$ and the identity or neutral element of the aggregation by $\overline{\theta}$. Pseudocode for the Recalculate-From-Scratch algorithm is shown in Algorithm \ref{pseudo:rfs}, assuming all functions have access to a \textit{vals} array which holds the window contents.

\begin{algorithm}
    \caption[Recalculate-From-Scratch]{Recalculate-From-Scratch insert, evict and query methods}
    \label{pseudo:rfs}
    \begin{algorithmic}[1]
        \Function{query}{}
            \State $agg\gets\overline{\theta}$
            \ForEach {$v \in vals$}
                \State $agg \gets agg \oplus v$
            \EndFor
            \State \Return $agg$
        \EndFunction

        \Function{insert}{v}
            \State $vals.pushBack(v)$
        \EndFunction

        \Function{evict}{}
            \State $vals.popFront()$
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element inserted or evicted from the window, the aggregation value is updated without recomputing the entire aggregation like in RFS. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations, that is, a function $\ominus$ exists such that $(x \oplus y) \ominus y = x$, for all $x$ and $y$. To illustrate the differences between SOE and RFS, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the current total sum plus ($\oplus$) the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed thus far minus ($\ominus$) the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable. Algorithm \ref{pseudo:soe} shows pseudocode for SOE assuming all functions have access to the window contents through the \textit{vals} array and access to the current aggregation state \textit{agg}.

\begin{algorithm}
    \caption[Subtract-On-Evict]{Subtract-On-Evict insert, evict and query methods}
    \label{pseudo:soe}
    \begin{algorithmic}[1]
        \Function{query}{}
            \State \Return $agg$
        \EndFunction

        \Function{insert}{v}
            \State $vals.pushBack(v)$
            \State $agg \gets agg \oplus v$
        \EndFunction

        \Function{evict}{}
            \State $v \gets vals.popFront()$
            \State $agg \gets agg \ominus val$
        \EndFunction
    \end{algorithmic}
\end{algorithm}

### 2020-07-05 chapters/sota.tex
Abbrevs

\subsection{Two-Stacks} \label{sec:2stacks}
The Two-Stacks SWAG algorithm works for \textit{associative} aggregations (recall Table \ref{tbl:aggregations-properties}) in scenarios where the sliding window has First-In First-Out (FIFO) semantics --- \textit{i.e.}, elements are evicted in the same order they were inserted. Two-Stacks \cite{Tangwongsan-DABA} implements the FIFO queue using two stacks, namely the front stack F and the back stack B. Each element in each of the stacks is a pair \textit{<val, agg>} where \textit{val} is the item's value and \textit{agg} is the aggregation value of everything below it on the stack. The pseudocode for Two-Stacks can be seen in Algorithm \ref{pseudo:2-stacks}, with $\oplus$ as the \textit{combine} function and $\overline{\theta}$ as the identity value of the \textit{combine} function, such that $\overline{\theta} \oplus x$ = $x \oplus \overline{\theta}$ = x. For example, if $\oplus$ is the arithmetic sum then $\overline{\theta}$ is 0. If $\oplus$ is the arithmetic multiplication then $\overline{\theta}$ is 1.

@@

        \Function{query}{F, B}
            \State \Return{getStackAggregationValue(F) $\oplus$ getStackAggregationValue(B)}
        \EndFunction

        \Function{insert}{B, v}
            \State{B.push( \{val: v, agg: getStackAggregationValue(B) $\oplus$ v\} }
        \EndFunction

        \Function{evict}{F, B}

            \If{F.isEmpty()} \Comment{// Flip B onto F and update top aggregation value}
                \While {\textbf{not} B.isEmpty()}
                    \State{F.push( \{val: B.top().val, agg: B.top().val $\oplus$ getStackAggregationValue(F) \})}
                    \State{B.pop()}
                \EndWhile
            \EndIf
            \State{F.pop()}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

The stochastic averaging process of HLL remains unchanged: incoming items are hashed and distributed across \textit{m} buckets. However, the authors propose keeping a short list of \textit{packets} per bucket instead of a single integer count. A \textit{packet} is defined as a pair <\textit{t\textsubscript{i}, p\textsubscript{i}}> where \textit{t\textsubscript{i}} is the arrival timestamp of the packet and \textit{p\textsubscript{i}} the position of the leftmost 1-bit in the binary hash representation of the associated value. Packets are only kept in the list if they are possible maxima over future windows. Because of that, the authors name this list a \textit{"List of Future Possible Maxima (LFPM)"}.

>>>>

\subsection{Two-Stacks} \label{sec:2stacks}
The Two-Stacks SWAG algorithm works for \textit{associative} aggregations (recall Table \ref{tbl:aggregations-properties}) in scenarios where the sliding window has First-In First-Out (FIFO) semantics --- \textit{i.e.}, elements are evicted in the same order they were inserted. Two-Stacks \cite{Tangwongsan-DABA} implements the FIFO queue using two stacks, namely the front stack F and the back stack B. Each element in each of the stacks is a pair \textit{<val, agg>} where \textit{val} is the item's value and \textit{agg} is the aggregation value of everything below it on the stack. The pseudocode for Two-Stacks can be seen in Algorithm \ref{pseudo:2-stacks}, assuming all functions have access to stacks F and B, with $\oplus$ as the \textit{combine} function and $\overline{\theta}$ as the identity value of the \textit{combine} function, such that $\overline{\theta} \oplus x$ = $x \oplus \overline{\theta}$ = x. For example, if $\oplus$ is the arithmetic sum then $\overline{\theta}$ is 0. If $\oplus$ is the arithmetic multiplication then $\overline{\theta}$ is 1.

@@

        \Function{query}{}
            \State \Return{getStackAggregationValue(F) $\oplus$ getStackAggregationValue(B)}
        \EndFunction

        \Function{insert}{v}
            \State{B.push( \{val: v, agg: getStackAggregationValue(B) $\oplus$ v\} }
        \EndFunction

        \Function{evict}{}
            \If{F.isEmpty()} \Comment{// Flip B onto F and update top aggregation value}
                \While {\textbf{not} B.isEmpty()}
                    \State{F.push( \{val: B.top().val, agg: B.top().val $\oplus$ getStackAggregationValue(F) \})}
                    \State{B.pop()}
                \EndWhile
            \EndIf
            \State{F.pop()}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

The stochastic averaging process of HLL remains unchanged: incoming items are hashed and distributed across \textit{m} buckets. However, the authors propose keeping a short list of \textit{packets} per bucket instead of a single integer count. A \textit{packet} is defined as a pair <\textit{t\textsubscript{i}, p\textsubscript{i}}> where \textit{t\textsubscript{i}} is the arrival timestamp of the packet and \textit{p\textsubscript{i}} the position of the leftmost 1-bit in the binary hash representation of the associated value. Packets are only kept in the list if they are possible maxima over future windows. Because of that, the authors name this list a ``List of Future Possible Maxima (LFPM)''.
